The \dealii Library, Version 9.7

### 2025-08-10 9.7/paper.tex
Small changes

\author[11]{Peter Munch}
 \affil[11]{Uppsala University, Sweden.
  {\texttt{peter.munch@it.uu.se}}}

@@

In this release, several extensions, major bug fixes and optimizations of the
matrix-free infrastructure and related multigrid functionality, such as
matrix-free geometric and polynomial coarsening, in \dealii have been
made. The most significant advances were made for the \kokkos-based
``portable'' infrastructure described below. In addition, improvements where
made to several areas:
\begin{itemize}
\item the \texttt{MatrixFreeTools::compute\_matrix()} and
  \texttt{MatrixFreeTools::compute\_diagonal()} functions to compute the
  matrix or only its diagonal from a matrix-free operator were enhanced
  towards multi-component systems,
\item support for additional ghost cells in the matrix-free setup was added to
  support vector access to a wider range operations, such as MPI
  parallelization of multigrid smoothers based on vertex
  patches~\cite{Wichrowski2025smoothers},
\item functionality for identifying degrees of freedom in lexicographic
  ordering on patches of cells,
\item support for additional matrices in the
  \texttt{TensorProductMatrixCreator} for support in fast
  diagonalization-based inverses.
\end{itemize}

@@

@@

Many of the \dealii tutorial programs were revised in a variety of ways
as part of this release: Around 125 of the more than 1500 (non-merge)
commits that went into this release touched the tutorial.
% data generated using these commands:
% - tutorial: git log --since 2024/08/11 --until 2025/07/19 --no-merges examples | grep commit | wc -l
% - total:    git log --since 2024/08/11 --until 2025/07/19 --no-merges | grep commit | wc -l
In addition, there are a number of new tutorial
programs:
\begin{itemize}
  \item
    \step{93}
    demonstrates how one can implement problems in which some of the
    variables to be solved for are not degrees of freedom of a finite
    element field (so-called ``non-local'' degrees of freedom, because
    they have no associated node functional that is tied to a cell or
    a patch of cells). Specifically, the program solves an
    optimization problem in which some of the variables are scalars
    that multiply certain source terms.
    \step{93} was written by Sam Scheuerman.
  \item
    \step{95} illustrates the use of matrix-free methods to solve
    problems with embedded boundaries using the CutFEM method that was
    previously already introduced in \step{85}. \step{95} was written
    by Maximilian Bergbauer, with help by Martin Kronbichler and Peter Munch.
  \item
    \step{97}
    is a program that solves a curl-curl problem from electromagnetics
    -- more specifically, from magnetostatics. The program shows a
    formulation in which two curl-curl problems are solved
    consecutively to model the magnetic field that results from the
    current in an electrically conducting coil.
    \step{97} was written by Siarhei Uzunbajakau.
\end{itemize}

In addition, there is one new program in the code gallery (a collection of
user-contributed programs that often solve more complicated problems
than tutorial programs, and that are intended as starting points for further
research rather than as teaching tools):
\begin{itemize}
  \item \emph{``Phase field fracture model in 3D''},
    contributed by Wasim Niyaz Munshi,
    Chandrasekhar Annavarapu,
    Wolfgang Bangerth, and
    Marc Fehling.
\end{itemize}

@@

P.~Munch acknowledges the funding by the Swedish Research Council (VR) under grant 2021-04620 and the strategic funding from the IT department of Uppsala University.

>>>>

\author[11]{Peter Munch}
 \affil[11]{Institute of Mathematics, Technical University of Berlin, Germany.
  {\texttt{muench@math.tu-berlin.de}}}

@@

In this release, several extensions, major bug fixes and optimizations of the
matrix-free infrastructure and related multigrid functionality, such as
matrix-free geometric and polynomial coarsening, in \dealii have been
made. The most significant advances were made for the \kokkos-based
``portable'' infrastructure described below. In addition, improvements where
made to several areas:
\begin{itemize}
\item the \texttt{MatrixFreeTools::compute\_matrix()} and
  \texttt{MatrixFreeTools::compute\_diagonal()} functions to compute the
  matrix or only its diagonal from a matrix-free operator were enhanced
  towards multi-component systems, the DG context (and, in general, to operators
  with face integrals), and mixed elements,
\item support for additional ghost cells in the matrix-free setup was added to
  support vector access to a wider range operations, such as MPI
  parallelization of multigrid smoothers based on vertex
  patches~\cite{Wichrowski2025smoothers},
\item functionality for identifying degrees of freedom in lexicographic
  ordering on patches of cells,
\item support for additional matrices in the
  \texttt{TensorProductMatrixCreator} for support in fast
  diagonalization-based inverses.
\end{itemize}

@@

\todo[inline]{Peter @ Luca: I think a code snippet would be great here! }

@@

Many of the \dealii tutorial programs were revised in a variety of ways
as part of this release: Around 125 of the more than 1500 (non-merge)
commits that went into this release touched the tutorial.
% data generated using these commands:
% - tutorial: git log --since 2024/08/11 --until 2025/07/19 --no-merges examples | grep commit | wc -l
% - total:    git log --since 2024/08/11 --until 2025/07/19 --no-merges | grep commit | wc -l
In addition, there are a number of new tutorial
programs:
\begin{itemize}
  \item
    \step{93}
    demonstrates how one can implement problems in which some of the
    variables to be solved for are not degrees of freedom of a finite
    element field (so-called ``non-local'' degrees of freedom, because
    they have no associated node functional that is tied to a cell or
    a patch of cells). Specifically, the program solves an
    optimization problem in which some of the variables are scalars
    that multiply certain source terms.
    \step{93} was written by Sam Scheuerman.
  \item
    \step{95} extends \step{85} and illustrates the use of matrix-free methods to solve
    problems with embedded boundaries using the CutFEM method~\cite{bergbauer2025high}. \step{95} was written
    by Maximilian Bergbauer, with help by Martin Kronbichler and Peter Munch.
  \item
    \step{97}
    is a program that solves a curl-curl problem from electromagnetics
    -- more specifically, from magnetostatics. The program shows a
    formulation in which two curl-curl problems are solved
    consecutively to model the magnetic field that results from the
    current in an electrically conducting coil.
    \step{97} was written by Siarhei Uzunbajakau.
\end{itemize}

In addition, there is one new program in the code gallery (a collection of
user-contributed programs that often solve more complicated problems
than tutorial programs, and that are intended as starting points for further
research rather than as teaching tools):
\begin{itemize}
  \item \emph{``Phase field fracture model in 3D''},
    contributed by Wasim Niyaz Munshi,
    Chandrasekhar Annavarapu,
    Wolfgang Bangerth, and
    Marc Fehling.

\todo[inline]{Peter: A short description would be great here!}
\end{itemize}

@@
